---
title: "survival_analysis_MM_xinyue"
author: "xinyue"
date: "9/15/2022"
output: html_document
---

```{r setup, include=FALSE}
#load necessary packages
fun_packages <- function(x){
  for( i in x ){
    if( ! require( i , character.only = TRUE ) ){
      #Si le package ne peut pas être chargé, on l'installe
      install.packages( i , dependencies = TRUE )
      #Charger le package après l'installation
      require( i , character.only = TRUE )
    }
    if (!requireNamespace("BiocManager", quietly = TRUE))
    {install.packages("BiocManager")}
  }
}

fun_packages(c("tidyverse","survival","survminer","devtools","org.Hs.eg.db","clusterProfiler","GGally","rms"))

#install.packages("devtools")
#BiocManager::install("survival")
#BiocManager::install("survminer")
#BiocManager::install("GGally")
#BiocManager::install("rms")
```

## Load clinical infomations and survival analysis with basical item: gender and age


```{r}
#load clinical informations
dat=read.csv2("sc3_Training_ClinAnnotations.csv",header=T,sep=",")
dat<-dat[,-which(apply(dat,2,function(x) all(is.na(x))))]

## with gender
dat$STATUS= ifelse(dat$HR_FLAG == 'TRUE', 1, 0)
my.surv <- Surv(dat$D_OS,dat$STATUS) 
kmfit1 <- survfit(my.surv~dat$D_Gender) 
ggsurvplot(kmfit1,data = dat)

#with age
my.surv_age <- Surv(dat$D_OS,dat$STATUS==1) 
kmfit2 <- survfit(my.surv_age~dat$D_Age) 
ggsurvplot(kmfit2,data = dat)

```

It seems that the gender doesn't matter if the survival time is less than 800 days, 
but once the patient could survival more than 800 days, female is likely to have more chance to live.

Elder people is more likely to die soon or have less probabilite to live after 800 days for the people more than 65 years old.

## survival analysis for each observed gene

The gene are selected by the model of RandomForest and performance of public scRNAseq data

Here is the list of gene that we chose:


```{r }
##############################  #########################
#observer one specific gene, for example here is :
## SNORD89  AURKAIP1  RNVU1.18   RNA5S10   RNA5S16  RNU1.1    RNA5S7  ELANE  LINC02108   RNA5S17
df_preprocess=read.csv2(file = 'df_preprocessing_and_merged.csv',header=T,sep=",")
df_preprocess=data.frame(df_preprocess)
Correct_Colnames <- function(df) {
  
  delete.columns <- grep("(^X$)|(^X\\.)(\\d+)($)", colnames(df), perl=T)
  
  if (length(delete.columns) > 0) {
    
    row.names(df) <- as.character(df[, grep("^X$", colnames(df))])
    #other data types might apply than character or 
    #introduction of a new separate column might be suitable
    
    df <- df[,-delete.columns]
    
    colnames(df) <- gsub("^X", "",  colnames(df))
    #X might be replaced by different characters, instead of being deleted
  }
  
  return(df)
}

df_select=Correct_Colnames(df_preprocess)

df_withoutHR=df_select[,-which(names(df_select)%in%c("HR_FLAG"))]

print("transform the entrez id to gene symbol")
TransSymbol<-function(df_select){
  trans_df=t(df_select)
  rank_names <- bitr(rownames(trans_df), 
                     fromType="ENTREZID", 
                     toType=c("SYMBOL"),
                     OrgDb="org.Hs.eg.db")
  expression_filter=data.frame(trans_df[rank_names$ENTREZID,])
  expression_filter$SYMBOL=rank_names$SYMBOL
  merge_data=merge(rank_names,expression_filter,by="SYMBOL",all=F)
  rownames(merge_data)=make.names(merge_data$SYMBOL,unique=T)
  data_pret=merge_data[ , -which(colnames(merge_data) %in% c("SYMBOL","ENTREZID"))]
  df_pret=data.frame(t(data_pret))
  return(df_pret)
}
df_select=TransSymbol(df_withoutHR)

rownames(dat)=dat$RNASeq_geneLevelExpFileSamplId
df_select <- mutate_all(df_select, function(x) as.numeric(as.character(x)))
df_merge=merge(df_select,dat, by="row.names",all.x=F)

```

## The application of previous selected gene list 

This gene list is chosen by the RandomForest and the typical biomarkers found at bibliography and then we did the observation in scRNAseq to see their distribution.

The list is: AURKAIP1,ANKRD11, NCOA3, CYLD, FAM46C,PRDM1,IRF4,ACTG1,

But through the previous test with our data, FAM46C doesn't exist in our case.

```{r }
my.surv <- Surv(dat$D_OS,dat$STATUS)
gene_list=c("AURKAIP1","ANKRD11", "NCOA3", "CYLD", "PRDM1","IRF4","ACTG1")

dat$AURKAIP1_group=ifelse(df_merge$AURKAIP1 > median(df_merge$AURKAIP1),'high','low')
dat$ANKRD11_group=ifelse(df_merge$ANKRD11 > median(df_merge$ANKRD11),'high','low')
dat$CYLD_group=ifelse(df_merge$CYLD > median(df_merge$CYLD),'high','low')
dat$PRDM1_group=ifelse(df_merge$PRDM1 > median(df_merge$PRDM1),'high','low')
dat$NCOA3_group=ifelse(df_merge$NCOA3 > median(df_merge$NCOA3),'high','low')
dat$IRF4_group=ifelse(df_merge$IRF4 > median(df_merge$IRF4),'high','low')
dat$ACTG1_group=ifelse(df_merge$ACTG1 > median(df_merge$ACTG1),'high','low')


kmfit1 <- survfit(my.surv~AURKAIP1_group,data = dat)
kmfit2<- survfit(my.surv~ANKRD11_group,data = dat)
kmfit3<- survfit(my.surv~CYLD_group,data = dat)
kmfit5<- survfit(my.surv~PRDM1_group,data = dat)
kmfit6<- survfit(my.surv~NCOA3_group,data = dat)
kmfit7<- survfit(my.surv~IRF4_group,data = dat)
kmfit8<- survfit(my.surv~ACTG1_group,data = dat)

ggsurvplot(kmfit1,conf.int =F, pval = T,risk.table =T, ncensor.plot = TRUE)
ggsurvplot(kmfit2,conf.int =F, pval = T,risk.table =T, ncensor.plot = TRUE)
ggsurvplot(kmfit3,conf.int =F, pval = T,risk.table =T, ncensor.plot = TRUE)
ggsurvplot(kmfit5,conf.int =F, pval = T,risk.table =T, ncensor.plot = TRUE)
ggsurvplot(kmfit6,conf.int =F, pval = T,risk.table =T, ncensor.plot = TRUE)
ggsurvplot(kmfit7,conf.int =F, pval = T,risk.table =T, ncensor.plot = TRUE)
ggsurvplot(kmfit8,conf.int =F, pval = T,risk.table =T, ncensor.plot = TRUE)

```

By observing those plots, it seems that there are AURKAIP1,ANKRD11 which are have more significant differences 

## Cox Proportional-Hazards Model

The math model is applied for our forest plot. 

``` {r}
################## Cox Proportional-Hazards Model #######################
formula <- formula(paste('Surv(D_OS, STATUS)~', gene_list))
summary(coxph(formula, data = df_merge))


uni_cox <- function(single_gene){
  formula <- as.formula(paste0('Surv(D_OS, STATUS)~', single_gene))
  surv_uni_cox <- summary(coxph(formula, data = df_merge))
  ph_hypothesis_p <- cox.zph(coxph(formula, data = df_merge))$table[1,3]
  if (surv_uni_cox$coefficients[,5]<0.05 & ph_hypothesis_p>0.05){  #get the pvalue
    single_cox_report <- data.frame('uni_cox_sig_genes'=single_gene,
                                    'beta'=surv_uni_cox$coefficients[,1],
                                    'Hazard_Ratio'=exp(surv_uni_cox$coefficients[,1]),
                                    'z_pvalue'=surv_uni_cox$coefficients[,5],
                                    'Wald_pvalue'=as.numeric(surv_uni_cox$waldtest[3]),
                                    'Likelihood_pvalue'=as.numeric(surv_uni_cox$logtest[3]))
    single_cox_report
  }
}
uni_cox_list <- lapply(gene_list, uni_cox)
candidate_gene=do.call(rbind, uni_cox_list)


#########################  modele construit ##################
#perform the multi-variates cox regression using qualified genes.
formula_for_multivariate <- as.formula(paste0('Surv(D_OS, STATUS)~', paste(candidate_gene$uni_cox_sig_genes, sep = '', collapse = '+')))
multi_variate_cox <- coxph(formula_for_multivariate, data = df_merge)
#check if variances are supported by PH hypothesis.
ph_hypo_multi <- cox.zph(multi_variate_cox)
#The last row of the table records the test results on the GLOBAL model. Delete it.
ph_hypo_table <- ph_hypo_multi$table[-nrow(ph_hypo_multi$table),]
#Remove variances not supported by ph hypothesis and perform the 2nd regression.
formula_for_multivariate <- as.formula(paste0('Surv(D_OS, STATUS)~', paste(rownames(ph_hypo_table)[ph_hypo_table[,3]>0.05], sep = '', collapse = '+')))
multi_variate_cox_2 <- coxph(formula_for_multivariate, data = df_merge)
```

## Forest plot
``` {r}
################ forest plot #####################
#observation of correlation of data
correlation <- cor(df_merge[,rownames(ph_hypo_table)[ph_hypo_table[,3]>0.05]], method = 'pearson')
ggpairs(df_merge[,rownames(ph_hypo_table)[ph_hypo_table[,3]>0.05]], 
        axisLabels = 'show')+
  theme_bw()+
  theme(panel.background = element_rect(colour = 'black', size=1, fill = 'white'), 
        panel.grid = element_blank())

vif <- rms::vif(multi_variate_cox_2)
#Some people said if the square root of VIF >2, they might be co-linear.
sqrt(vif) < 2

ggforest(model = multi_variate_cox_2, data = df_merge, main = 'Hazard ratios of candidate genes', fontsize = 1)

```
